Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}

### 2020-07-21 llncs2e/paper12.tex
typo
paper12

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{euler}
\usepackage{amsmath,amssymb,amscd,amstext}
\usepackage{xcolor}
\usepackage{pst-node}
\usepackage{tikz,pgf}
\usepackage{tikz-cd} 
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usetikzlibrary{spy,decorations.markings,decorations.text,arrows.meta,bending,3d,shapes.arrows,positioning,fit,backgrounds,fadings}
\let\clipbox\relax
\usepackage{slashbox}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{multibib}
\usepackage{filecontents}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{commath}
\usepackage[percent]{overpic}
\usepackage[hidelinks]{hyperref}
\hypersetup{
    pdftitle={Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory},    % title
    pdfauthor={Luciano Melodia},     % author
    pdfsubject={Estimate of the Neural Network Dimension},   % subject of the document
    pdfcreator={Luciano Melodia},   % creator of the document
    pdfproducer={Luciano Melodia},  % producer of the document
    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {neural networks} {Manifold learning}, % list of keywords
    pdfnewwindow=true,
    colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=purple
}
\definecolor{red1}{RGB}{227,226,223}
\definecolor{red2}{RGB}{227,175,188}
\definecolor{red3}{RGB}{238,76,124}
\definecolor{red4}{RGB}{154,23,80}
\definecolor{red5}{RGB}{93,0,30}
\makeatletter
\newcommand{\@chapapp}{\relax}
\newcommand{\xdashrightarrow}[2][]{\ext@arrow 0359\rightarrowfill@@{#1}{#2}}
\newcommand{\xdashleftarrow}[2][]{\ext@arrow 3095\leftarrowfill@@{#1}{#2}}
\newcommand{\xdashleftrightarrow}[2][]{\ext@arrow 3359\leftrightarrowfill@@{#1}{#2}}
\def\rightarrowfill@@{\arrowfill@@\relax\relbar\rightarrow}
\def\leftarrowfill@@{\arrowfill@@\leftarrow\relbar\relax}
\def\leftrightarrowfill@@{\arrowfill@@\leftarrow\relbar\rightarrow}
\def\arrowfill@@#1#2#3#4{%
  $\m@th\thickmuskip0mu\medmuskip\thickmuskip\thinmuskip\thickmuskip
   \relax#4#1
   \xleaders\hbox{$#4#2$}\hfill
   #3$%
}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\makeatother
\begin{document}
%
\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\titlerunning{Estimate of the neural network Dimension}

@@

\section{Experimental setting}
\input{figure2.tex}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The layer size is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural networks around a hundred times. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.\\[0.4cm]
\textbf{Persistence Landscapes:} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
\textbf{Neural networks:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural networks are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$ for the purpose of fitting into a dense layer. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to stay on the same differentiable structure during learning. A total of $5$ hidden layers are used, all with the same embedding dimension. The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow v.2.2.0} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]
\textbf{Results:} In Fig. \ref{figure2} c) d) we colored the neural networks according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.  The embedding dimension can be taken from Tab. \ref{figure3}.\footnote{Due to the construction of invertibility the treshold must be doubled, see \cite{DinhSB17}.} All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig. \ref{figure2} c) d). The models above the treshold remain stable. A single net still shows an explosive gradient colored ${\color{red4}\bullet}$ in Fig. \ref{figure2} d). The net showing the deflection of gradient is very close to our threshold, identified as $292$ dimensional embedding. The closer we come to the treshold, the sparser the fluctuations of loss.
\input{figure3.tex}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{euler}
\usepackage{amsmath,amssymb,amscd,amstext}
\usepackage{xcolor}
\usepackage{pst-node}
\usepackage{tikz,pgf}
\usepackage{tikz-cd} 
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usetikzlibrary{spy,decorations.markings,decorations.text,arrows.meta,bending,3d,shapes.arrows,positioning,fit,backgrounds,fadings}
\let\clipbox\relax
\usepackage{slashbox}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{multibib}
\usepackage{filecontents}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{commath}
\usepackage[percent]{overpic}
\usepackage[hidelinks]{hyperref}
\hypersetup{
    pdftitle={Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory},    % title
    pdfauthor={Luciano Melodia},     % author
    pdfsubject={Estimate of the Neural Network Dimension},   % subject of the document
    pdfcreator={Luciano Melodia},   % creator of the document
    pdfproducer={Luciano Melodia},  % producer of the document
    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {neural networks} {Manifold learning}, % list of keywords
    pdfnewwindow=true,
    colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=blue
}
\definecolor{red1}{RGB}{227,226,223}
\definecolor{red2}{RGB}{227,175,188}
\definecolor{red3}{RGB}{238,76,124}
\definecolor{red4}{RGB}{154,23,80}
\definecolor{red5}{RGB}{93,0,30}
\makeatletter
\newcommand{\@chapapp}{\relax}
\newcommand{\xdashrightarrow}[2][]{\ext@arrow 0359\rightarrowfill@@{#1}{#2}}
\newcommand{\xdashleftarrow}[2][]{\ext@arrow 3095\leftarrowfill@@{#1}{#2}}
\newcommand{\xdashleftrightarrow}[2][]{\ext@arrow 3359\leftrightarrowfill@@{#1}{#2}}
\def\rightarrowfill@@{\arrowfill@@\relax\relbar\rightarrow}
\def\leftarrowfill@@{\arrowfill@@\leftarrow\relbar\relax}
\def\leftrightarrowfill@@{\arrowfill@@\leftarrow\relbar\rightarrow}
\def\arrowfill@@#1#2#3#4{%
  $\m@th\thickmuskip0mu\medmuskip\thickmuskip\thinmuskip\thickmuskip
   \relax#4#1
   \xleaders\hbox{$#4#2$}\hfill
   #3$%
}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\makeatother
\begin{document}
%
\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\titlerunning{Estimate of the neural network Dimension}

@@

\section{Experimental setting}
\input{figure2.tex}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The layer size is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural networks have been trained around a hundred times each. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.\\[0.4cm]
\textbf{Persistence Landscapes:} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
\textbf{Neural networks:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural networks are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$ for the purpose of fitting into a dense layer. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to stay on the same differentiable structure during learning. A total of $5$ hidden layers are used, all with the same embedding dimension. The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow v.2.2.0} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]
\textbf{Results:} In Fig. \ref{figure2} c) d) we colored the neural networks according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.  The embedding dimension can be taken from Tab. \ref{figure3}.\footnote{Due to the construction of invertibility the treshold must be doubled, see \cite{DinhSB17}.} All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig. \ref{figure2} c) d). The models above the treshold remain stable. A single net still shows an explosive gradient colored ${\color{red4}\bullet}$ in Fig. \ref{figure2} d). The net showing the deflection of gradient is very close to our threshold, identified as $292$ dimensional embedding. The closer we come to the treshold, the sparser the fluctuations of loss.
\input{figure3.tex}
